Will Kamala Harris will win the 2024 Presidential election

### 2024-11-04 paper/paper.qmd
fixed refrences

library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
library(broom)
library(modelsummary)
library(rstanarm)
library(knitr)
```

@@

¿ does this go into data cleaning ?

We use the statistical programming language R [@citeR]. Our data comes from [@fivethirtyeight2024] and was cleaned, modeled and graphed, using @broom, @tidy, @rstanarm, @janitor, @lubr, @modelsummary.

@@

In this analysis, we examine polling trends for Kamala Harris across key swing states, focusing on recent polling data from Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin [CITE NYT PAPER]. As shown in @fig-supportstate, which displays the support percentages for both Harris and Donald Trump side-by-side by state, the data reveals a competitive race, with relatively close support levels for both candidates in most states. While Trump has a slight edge in Arizona, Georgia, and North Carolina, Harris maintains a modest lead in Michigan, Nevada, and Pennsylvania.

@@

Interestingly enough, if we look at @fig-supportGNN we can see Nevada, similar to several other states had a spike in support of Harris in early September peaking north of 48% in the polls and their recent history of voting in favor of the democratic candidates. Nevertheless, Harris is seen to have had a gradual decline in the state stabilizing at around 46%, which which seems low in comparison to other states that are looking more favorably at Harris than Trump. Still, when we look at @supportstate Harris is still has a marginal to Trump, considering all of these factors it is likely Harris will in fact take Nevada for the democrats.

@@

The New York Times/Siena College Poll is conducts live telephone interviews with a national sample of registered voters, using both landlines and cellphones, with over 90% of contacts made via cellphones. The sample frame is drawn from a comprehensive voter file covering 200 million registered voters. This file provides demographic and geographic information, enabling representative sampling across age, race, and political affiliation. The poll targets registered U.S. voters, with special focus on battleground states like Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin, determined as swing states. Respondents are randomly selected from the voter file, ensuring a diverse sample by region, party, and demographics. To mitigate non-response bias, the poll applies demographic weighting, increasing the weight of underrepresented groups, such as young adults or non-college graduates. Low response rates (around 2%) present challenges, to mitigate this, weighing adjustments are applied to correct for demographic imbalances and potential non-response bias. Questions are structured to be clear, neutral, and representative across political views, with each question carefully reviewed for balance and clarity. This process reduces the risk of bias in responses, though the brief format of phone surveys limits depth.

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-ppcheck
#| fig-cap: "Posterior predictive check of model"

@@

In @fig-ppcheck, we check the Posterior predictive check of model of our model. Since the raw data aligns closely with our prdictions, our model is a good predicton of the data.

>>>>

library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
library(modelsummary)
library(rstanarm)
library(knitr)
library(bayesplot)
```

@@

We use the statistical programming language R [@citeR] to compile this paper. Our data comes from [@fivethirtyeight2024] and was cleaned and visualized using the packages Tidyverse [@tidy], Lubridate [@lubr], Janitor [@janitor], and Knitr [@knitr] and modeled using Rstanarm [@rstanarm], ModelSummary [@modelsummary] and BayesPlot [@bayesplot].

@@

In this analysis, we examine polling trends for Kamala Harris across key swing states, focusing on recent polling data from Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin [@NYTswingatates]. As shown in @fig-supportstate, which displays the support percentages for both Harris and Donald Trump side-by-side by state, the data reveals a competitive race, with relatively close support levels for both candidates in most states. While Trump has a slight edge in Arizona, Georgia, and North Carolina, Harris maintains a modest lead in Michigan, Nevada, and Pennsylvania.

@@

Interestingly enough, if we look at @fig-supportGNN we can see Nevada, similar to several other states had a spike in support of Harris in early September peaking north of 48% in the polls and their recent history of voting in favor of the democratic candidates. Nevertheless, Harris is seen to have had a gradual decline in the state stabilizing at around 46%, which which seems low in comparison to other states that are looking more favorably at Harris than Trump. Still, when we look at @fig-supportstate Harris is still has a marginal to Trump, considering all of these factors it is likely Harris will in fact take Nevada for the democrats.

@@

The New York Times/Siena College Poll is conducts live telephone interviews with a national sample of registered voters, using both landlines and cellphones, with over 90% of contacts made via cellphones. The sample frame is drawn from a comprehensive voter file covering 200 million registered voters. This file provides demographic and geographic information, enabling representative sampling across age, race, and political affiliation. The poll targets registered U.S. voters, with special focus on battleground states like Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin, determined as swing states. Respondents are randomly selected from the voter file, ensuring a diverse sample by region, party, and demographics. To mitigate non-response bias, the poll applies demographic weighting, increasing the weight of underrepresented groups, such as young adults or non-college graduates. Low response rates (around 2%) present challenges, to mitigate this, weighing adjustments are applied to correct for demographic imbalances and potential non-response bias. Questions are structured to be clear, neutral, and representative across political views, with each question carefully reviewed for balance and clarity. This process reduces the risk of bias in responses, though the brief format of phone surveys limits depth (@NYTpolls).

@@

```{r, fig.width=5, fig.height=3}
#| echo: false
#| warning: false
#| message: false
#| label: fig-ppcheck
#| fig-cap: "Posterior predictive check of model"

@@

In @fig-ppcheck, we check the Posterior predictive check of model of our model. Since the raw data aligns closely with our predictions, our model is a good prediction of the data.
